---
title: "Group 13 Final Project"
author: "cs3779, kd2640, ob2305, mp3745, lef2147"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: html_document
---

Rough Outline of Project:

* Describe the goal/motivation of the project; have some stats and facts to explain our purpose in investigating this dataset
* Describe the dataset: where is it from, what does it contain
* Review the questions we intend on answering 
* Exploratory Data Analysis
  * Give an overview of the demographics of our dataset, any conclusion we make is only relevant to those who were stopped which is why it is good to give an idea of who is represented
    * histogram of race/build/etc.
    * distribution of continuous variables
  * Logistic Regression
    * odds of frisked
    * odds of arrest
  * Trends over time
    * could maybe find a way to test this, i.e. linear regression, is the slope equal to zero, or something of that nature. the plot shows a funnel pattern so maybe we could look into that more
  * other things
* Findings/Discussion




```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(rvest)
library(httr)
library(lubridate)
library(plotly)
library(rgdal)
library(sp)
library(sf)
library(leaflet)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


Read in and tidy the data

The following code:

* Reads in the data
* Renames columns to be more informative
* Combines height columns into a single height in inches
* Converts date_stop to date data type
* Converts time_stop to time data type
* Recodes the values in categorical columns to be more informative
* Selects column subset for further analysis

```{r, message=FALSE}
# Read in data
stop_frisk_df = 
  # Read in data from internet
  GET("https://www1.nyc.gov/assets/nypd/downloads/excel/analysis_and_planning/stop-question-frisk/sqf-2016.csv") %>% 
  content("parsed") %>% 
  
  # Clean and fix names of columns
  janitor::clean_names() %>% 
  rename(
    precinct = pct,
    date_stop = datestop,
    time_stop = timestop,
    stop_in_out = inout,
    obs_time_min = perobs,
    stop_time_min = perstop,
    arst_made = arstmade,
    off_in_unif = offunif,
    hair_col = haircolr,
    eye_col = eyecolor,
    other_feature = othfeatr,
    boro = city
  )  %>% 
  mutate(
    # Combine height columns
    height_inch = ht_feet * 12 + ht_inch,
    # Convert date to proper format
    date_stop = mdy(date_stop),
    # Convert time to proper format
    time_stop = hm(time_stop / 100),
    # Recode to be more informative
    stop_in_out = recode(stop_in_out, "I" = "inside", "O" = "outside"),
    race = recode(
      race, 
      "A" = "asian/pacific islander", 
      "B" = "black", 
      "I" = "american indian/alaska native",
      "P" = "black-hispanic",
      "Q" = "white-hispanic",
      "W" = "white",
      "U" = "unknown",
      "Z" = "other"
    ),
    hair_col = recode(
      hair_col,
      "BA" = "bald",
      "BK" = "black",
      "BL" = "blond",
      "BR" = "brown",
      "DY" = "dyed",
      "FR" = "frosted",
      "GY" = "gray",
      "RA" = "red",
      "SN" = "sandy",
      "SP" = "salt and pepper",
      "WH" = "white",
      "XX" = "unknown",
      "ZZ" = "other",
    ),
    eye_col = recode(
      eye_col,
      "BK" = "black",
      "BL" = "blue",
      "BR" = "brown",
      "DF" = "different",
      "GR" = "green",
      "GY" = "gray",
      "HA" = "hazel",
      "MA" = "maroon",
      "PK" = "pink",
      "VI" = "violet",
      "XX" = "unknown",
      "Z" = "other",      
    ),
    build = recode(
      build,
      "H" = "heavy",
      "M" = "medium",
      "T" = "thin",
      "U" = "muscular",
      "Z" = "unknown"
    ),
    # change boro columns to lowercase for consistency
    boro = tolower(boro),
    # change character datatypes to numeric
    age = as.numeric(age),
    obs_time_min = as.numeric(obs_time_min),
    stop_time_min = as.numeric(stop_time_min)
  )  %>% 
  # select columns for further analysis
  select(precinct, date_stop, time_stop, stop_in_out, obs_time_min, stop_time_min, arst_made, off_in_unif, frisked, 
         searched, rf_vcrim, rf_othsw, rf_attir:ac_evasv, cs_furtv:cs_other, rf_knowl, sb_hdobj:sb_admis, rf_furt, 
         rf_bulg, sex, race, age, height_inch, weight:build, boro, xcoord, ycoord) %>% 
  # change all columns that have Y/N to 1/0
  mutate_at(vars(arst_made:rf_bulg), funs(recode(., "Y" = "1", "N" = "0"))) %>% 
  # change binary columns to numeric instead of character
  mutate_at(vars(arst_made:rf_bulg), funs(as.numeric(.))) %>% 
  # converts all character variables to factors (this does the same as the for loop)
  mutate_if(is.character, as.factor) %>% 
  # remove the single row of NAs
  filter(!is.na(build))
```


Evaluating Missing Data and Categorical Data
```{r}
table(stop_frisk_df$race)
table(stop_frisk_df$hair_col)
table(stop_frisk_df$eye_col)
table(stop_frisk_df$build)
# we should consider consolidating categories

colSums(is.na(stop_frisk_df))
# we should consider removing variable other_feature (11637 missing obs)
# age has 35 missing values, consider multiple imputation methods here

# Looks like there is also an entire row of NA's
```


Looking at stops over time 

* Over a year
* By time of day

```{r, message=FALSE}
# Number of stops per day
stop_frisk_df %>% 
  group_by(date_stop) %>% 
  summarize(
    count = n()
  ) %>% 
  ggplot(aes(x = date_stop, y = count)) + 
  geom_point() +
  geom_smooth(se = FALSE)

# Number of stops per month
stop_frisk_df %>% 
  mutate(
    month_stop = factor(month(date_stop))
  ) %>% 
  filter(
    month_stop != is_null(month_stop)
  ) %>% 
  group_by(month_stop) %>% 
  summarize(
    count = n()
  ) %>% 
  ggplot(aes(x = month_stop, y = count)) + 
  geom_bar(stat = "Identity") 

# Number of stops per day (broken down by boro)
stop_frisk_df %>% 
  mutate(
    month_stop = month(date_stop)
  ) %>% 
  filter(
    month_stop != is_null(month_stop)
  ) %>%   
  group_by(month_stop, boro) %>% 
  summarize(
    count = n()
  ) %>% 
  ggplot(aes(x = month_stop, y = count, color = boro)) + 
  geom_point() +
  geom_smooth(se = FALSE) 

# Number of stops per hour over the day
stop_frisk_df %>% 
  mutate(
    hour_stop = hour(time_stop),
    part_of_day = cut(hour_stop, breaks = c(0, 5, 12, 17, 21, 24), labels = c("Night", "Morning", "Afternoon", "Evening", "Night"), right = FALSE)
  ) %>% 
  select(hour_stop, part_of_day) %>% 
  group_by(hour_stop, part_of_day) %>% 
  summarize(
    count = n()
  ) %>% 
  ggplot(aes(x = hour_stop, y = count)) + 
  geom_bar(stat = "Identity", aes(fill = part_of_day)) +
  geom_smooth(se = FALSE)

# Number of stops per hour over the day (broken down by boro)
stop_frisk_df %>% 
  mutate(
    hour_stop = hour(time_stop)
  ) %>% 
  group_by(hour_stop, boro) %>% 
  summarize(
    count = n()
  ) %>% 
  ggplot(aes(x = hour_stop, y = count, color = boro)) + 
  geom_point() +
  geom_smooth(se = FALSE)
```

This code chunk looks at the number of people stopped, frisked, frisked & searched, and searched over a single day

```{r, message=FALSE}
stop_frisk_df %>% 
  filter(date_stop != is.na(date_stop)) %>% 
  mutate(
    hour_stop = hour(time_stop)
  ) %>% 
  group_by(hour_stop, frisked, searched) %>% 
  summarize(
    count = n()
  ) %>% 
  pivot_wider(
    names_from = frisked:searched,
    values_from = count
  ) %>% 
  rename(
    "stopped" = "0_0",
    "frisked" = "1_0",
    "searched" = "0_1",
    "frisk_and_search" = "1_1"
  ) %>% 
  mutate(
    searched = replace_na(searched, 0),
    stopped = replace_na(stopped, 0),
    frisked = replace_na(frisked, 0),
    frisk_and_search = replace_na(frisk_and_search, 0),
    stopped = stopped + searched + frisked + frisk_and_search
  ) %>% 
  ggplot(aes(x = hour_stop, y = stopped)) +
  geom_smooth(se = FALSE, color = 'red') +
  geom_smooth(aes(y = frisked, color = 'blue'), se = FALSE) +
  geom_smooth(aes(y = searched, color = 'green'), se = FALSE) +
  geom_smooth(aes(y = frisk_and_search, color = 'yellow'), se = FALSE) 
```




















Logistic Regression Dataset
  - not sure if we want to use this for overall but didnt want to completely change the code above without asking everyone
  
```{r}
stop_frisk_log = stop_frisk_df %>% mutate(
    race = recode(
      race, 
      "asian/pacific islander" = "other", 
      "black" = "black", 
      "american indian/alaska native" = "other",
      "black-hispanic" = "black-hispanic",
      "white-hispanic" = "white-hispanic",
      "white" = "white",
      "unknown" = "other",
      "other" = "other"
    ),
    hair_col = recode(
      hair_col,
      "dyed" = "other",
      "frosted" = "other",
      "gray" = "other",
      "red" = "other",
      "sandy" = "other",
      "salt and pepper" = "other",
      "white" = "other",
      "unknown" = "other",
      "ZZ" = "other"
    ),
    eye_col = recode(
      eye_col,
      "different" = "other",
      "green" = "other",
      "gray" = "other",
      "hazel" = "other",
      "maroon" = "other",
      "pink" = "other",
      "violet" = "other",
      "unknown" = "other"
    ),
    build = recode(
      build,
      "muscular" = "other",
    ))
```






Building a model using only characteristics, demographics, and location as predictors for frisked
 - assess multicolinearity to determine if there are any variables that exhibit high correlation 
 - We will remove any variables that exhibit signs of multicollinearity 
 - From this model, we can assess of those who were stopped, what are the odds of getting frisked based on certain characteristics, demographics and location predictors
    - IMPORTANT: we cannot make any causal statements from this model 'i.e. you have a higher odds of getting stopped if you are X race' because everyone in this dataset was stopped
    
```{r}
model_1 = glm(frisked ~ sex + race + age + height_inch + weight + hair_col + eye_col + boro + build + stop_in_out + precinct + off_in_unif, family = binomial, data = stop_frisk_log)

car::vif(model_1)
# Based on the GVIF, we will remove boro

model_2 = glm(frisked ~ sex + race + age + height_inch + weight + hair_col + eye_col + build + stop_in_out + precinct + off_in_unif, family = binomial, data = stop_frisk_log)

car::vif(model_2)
# no more collinearity problems

summary(model_2)
# Remove hair and eye color because all categories within them are highly unsignificant 

model_3 = glm(frisked ~ sex + race + age + height_inch + weight + build + stop_in_out + precinct + off_in_unif, family = binomial, data = stop_frisk_log)

summary(model_3)
# keep weight? 
```

Building a model that uses characteristics, demographics, and location as predictors for arrest made
```{r}
model_4 = glm(arst_made ~ sex + race + age + height_inch + weight + hair_col + eye_col + build + stop_in_out + precinct + off_in_unif, family = binomial, data = stop_frisk_log)

car::vif(model_4)

summary(model_4)

# remove eye color
model_5 = glm(arst_made ~ sex + race + age + height_inch + weight + hair_col + build + stop_in_out + precinct + off_in_unif, family = binomial, data = stop_frisk_log)

summary(model_5)

# remove build
model_6 = glm(arst_made ~ sex + race + age + height_inch + weight + hair_col + stop_in_out + off_in_unif, family = binomial, data = stop_frisk_log)

summary(model_6)


# remove height
model_7 = glm(arst_made ~ sex + race + age + weight + hair_col + stop_in_out + off_in_unif, family = binomial, data = stop_frisk_log)

summary(model_7)
```


```{r}
stop_frisk_log %>% group_by(frisked, searched, arst_made) %>% 
  summarise(
    n_obs = n())
```


```{r, demographics, eval = FALSE}

# count per sex group
demographics = stop_frisk_log %>% 
  select(sex:build, height_inch, frisked, searched) 

# count subjects by sex
demographics %>% 
  group_by(sex) %>% 
  summarize(count = n()) %>% 
  knitr::kable()
  
# distribution of age by sex
demographics %>% 
  plot_ly(y = ~ age, color = ~sex, type = "violin", colors = "Set2")


# percentage of races stopped
demographics %>% 
  count(race) %>% 
  mutate(percent = (n / nrow(demographics))* 100,
         # percent = str_replace(percent, "[0-9]$", "%"),
         race = as.factor(race),
         race = fct_reorder(race, n)) %>% 
  plot_ly(x = ~ race, y = ~ percent, type = "bar", color = ~ race)

  

# subjects count of frisked subjects by race, sex 
frisked_plot = 
  stop_frisk_df %>% 
  drop_na() %>% 
  mutate(
    race = fct_infreq(race)
    ) %>% 
  filter( 
    frisked == '1') %>% 
  ggplot(aes(x = race, fill = sex))+
  geom_bar(alpha = .5, position = "dodge")+
  labs(
    title = "Frisked")+
  theme(axis.text.x = element_text(angle = 80, hjust = 1))

# subjects count of searched subjects by race, sex
searched_plot = 
  stop_frisk_df  %>% 
  drop_na() %>% 
  mutate(
    race = fct_infreq(race)) %>% 
  filter(searched =='1') %>% 
  ggplot(aes(x = race, fill = sex))+
  geom_bar(alpha = .5, position = "dodge")+
  labs(
    title = "Searched")+
  theme(axis.text.x = element_text(angle = 80, hjust = 1))

# subjects count of arrested subjects by race, sex
arrested_plot = 
  stop_frisk_df  %>% 
  drop_na() %>% 
  mutate(
    race = fct_infreq(race)) %>% 
  filter(arst_made =='1') %>% 
  ggplot(aes(x = race, fill = sex))+
  geom_bar(alpha = .5, position = "dodge")+
  labs(
    title = "Arrested")+
  theme(axis.text.x = element_text(angle = 80, hjust = 1))

# devtools::install_github("thomasp85/patchwork")
library(patchwork)

(frisked_plot+searched_plot + arrested_plot)

# final plot
stop_frisk_log %>% 
  drop_na() %>% 
  mutate(stopped = as.numeric(stop_in_out),
         stopped = case_when(
           stopped > 0 ~ '1',
           TRUE ~""),
         stopped = as.numeric(stopped)) %>% 
  select(race, arst_made, frisked, stopped) %>% 
  pivot_longer(
      arst_made:stopped,
      names_to = "action",
      values_to = "value") %>% 
  filter(value == '1') %>% 
  count(race, action) %>% 
  mutate(race = fct_infreq(race, n)) %>% 
  plot_ly(x = ~action, y = ~n, type = "bar", color = ~ race)

```

Regression model for demographics

```{r, eval=FALSE}
dem_model = glm(arst_made ~ sex:build + height_inch, family = binomial, data = stop_frisk_log)
summary(dem_model)
car::vif(dem_model)
# nothing is sig

arrest_model = glm(arst_made ~ sex + age + weight, family = binomial, data = stop_frisk_log)
summary(arrest_model)
__________________________
# did not want to change the whole dataset. Releveles arst_made, so regression models outcome arrested 

stop_frisk_relevel = 
  stop_frisk_log %>% 
  mutate(
    arst_made = as.factor(arst_made),
    arst_made = fct_relevel(arst_made, '1'))
  
arrest_model_2 = 
  glm(arst_made ~ sex + age + weight, family = binomial, data = stop_frisk_relevel)
summary(arrest_model_2)
```


Stops of subjects below 18. Maybe shny?
```{r, eval = FALSE}
# percent innocent under 18
under_18 = stop_frisk_log %>% 
  filter(age<'18') 

stop_frisk_log %>% 
  filter(age < '18') %>% 
  count(arst_made) %>%
  mutate(
    percent = (n /nrow(under_18)* 100)
  ) %>% 
  knitr::kable(digits = 2)
  
# frisked under 18 by race
stop_frisk_log %>% 
  filter(
    age < '18',
    frisked == '1') %>%
  count(race) %>% 
  mutate(
    race = fct_reorder(race, -n)) %>% 
  plot_ly(x = ~ race, y = ~ n, type = "bar", color = ~ race)


#searched under 18
stop_frisk_log %>% 
  filter(
    age < '18',
    searched == '1') %>%
  count(race) %>% 
  mutate(
    race = fct_reorder(race, -n)) %>% 
  plot_ly(x = ~ race, y = ~ n, type = "bar", color = ~ race)


#arrested under 18
stop_frisk_log %>% 
  filter(
    age < '18',
    arst_made =='1') %>%
  count(race) %>% 
  mutate(
    race = fct_reorder(race, -n)) %>% 
  plot_ly(x = ~ race, y = ~ n, type = "bar", color = ~ race)


# stoped under 18
stop_frisk_log %>% 
  filter(
    age < '18') %>%
  count(race) %>% 
  mutate(
    race = fct_reorder(race, -n)) %>% 
  plot_ly(x = ~ race, y = ~ n, type = "bar", color = ~ race)

# innocent under 18
stop_frisk_log %>% 
  filter(
    age < '18',
    arst_made == '0') %>%
  count(race) %>% 
  mutate(
    race = fct_reorder(race, -n)) %>% 
  plot_ly(x = ~ race, y = ~ n, type = "bar", color = ~ race)

# percent innocent under 18
stop_frisk_log %>% 
  filter(
    age < '18',
    arst_made =='0') %>% 
  count(race) %>% 
  mutate(
    percent = (n/nrow(under_18)* 100),
    race = fct_reorder(race, -n)) %>% 
  plot_ly(x = ~ race, y = ~ percent, type = "bar", color = ~ race)

# reasons for stoping under 18
stop_frisk_df %>% 
  filter(age < '18') %>%
  select(cs_objcs:cs_lkout, cs_cloth, cs_drgtr, cs_furtv, cs_vcrim:cs_other) %>% 
  pivot_longer(
    cs_objcs:cs_other,
    names_to = "reason_stopped",
    values_to = "stops"
  ) %>% 
  mutate(
    reason_stopped = recode(
    reason_stopped,
    "cs_objcs" = "carrying suspicious object",
    "cs_descr" = "fits a relevant description",
    "cs_casng" = "casing a victim or location",
    "cs_lkout" = "suspect acting as a lookout",
    "cs_cloth" = "wearing clothes commonly used in crimes",
    "cs_drgtr" = "actions indicative of drug transaction",
    "cs_furtv" = "furtive movements",
    "cs_vcrim" = "actions engaging in violent crime",
    "cs_bulge" = "suspcious bulge",
    "cs_other" = "other"
  )) %>% 
  filter(stops == 1) %>%
  group_by(reason_stopped) %>% 
  summarize(total =n()) %>% 
  mutate(prob = total/sum(total)) %>% 
  plot_ly(x = ~reason_stopped, y= ~total, type = "bar" , color = ~reason_stopped)

# percent innocent in black race category
stop_frisk_log %>% 
  filter(age < 18,
         race == 'black') %>% 
  group_by(race, arst_made) %>% 
  count() %>% 
  pivot_wider(
    names_from = arst_made,
    values_from = n) %>% 
  rename(
    innocent = `0`,
    arrested = `1`) %>% 
  mutate(
    percentage = (innocent/(innocent + arrested)) *100)



# function to do the same for every race
under_18_inn = function(x) {
  stop_frisk_log %>% 
  filter(age < 18,
         race == x) %>% 
  group_by(race, arst_made) %>% 
  count() %>% 
  pivot_wider(
    names_from = arst_made,
    values_from = n) %>% 
  rename(
    innocent = `0`,
    arrested = `1`) %>% 
  mutate(
    percentage = (innocent/(innocent + arrested)) *100)
}

output = map_df(stop_frisk_df$race, under_18_inn) %>% 
  distinct(race, percentage)

output %>% 
  ungroup(race) %>% 
  mutate(
    race = fct_relevel(race, c("white", "black", "other", "white-hispanic")),
    race = recode(race, 
                  "white" = "White",
                  "black" = "Black",
                  "other" = "Other",
                  "white-hispanic" = "White-Hispanic",
                  "black-hispanic" = "Black-Hispanic")
    ) %>% 
plot_ly(x = ~race, y = ~percentage, type = "bar", color = ~race) %>% 
  layout(
    title = 'Percent innocent Minors amoung each race category',
    xaxis = list(
      type = 'category',
      title  = 'Race'), 
    yaxis = list(
      title = 'Percent innocent, %',
      range = c(0, 100)
    ) 
  )

```



## Converting X Y Coordinates
```{r}
test_df= 
  stop_frisk_df %>% 
  select(xcoord, ycoord) %>% 
  drop_na()

coordinates(test_df) <- c("xcoord","ycoord")
proj4string(test_df) <- CRS("+init=epsg:2263")
CRS.new <- CRS("+init=epsg:4326") 
test_df_new <- spTransform(test_df, CRS.new)
test_df_new <- data.frame(longitude = coordinates(test_df_new)[,1], latitude = coordinates(test_df_new)[,2])
```

# Simplifying code
```{r}
long_lat_df =
  stop_frisk_df %>% 
  select(xcoord, ycoord) %>% 
  drop_na()

coordinates(long_lat_df) <- c("xcoord", "ycoord")
proj4string(long_lat_df) <- CRS("+init=epsg:2263")
long_lat_df <- spTransform(long_lat_df, CRS("+init=epsg:4326"))
long_lat_df <- data.frame(longitude = coordinates(long_lat_df)[,1], latitude = coordinates(long_lat_df)[,2])
```














