---
title: "mp3745_time_analysis"
author: "mp3745"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(rvest)
library(httr)
library(lubridate)
library(gganimate)
library(transformr)
library(gifski)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

Read in and tidy the data

The following code:

* Reads in the data
* Renames columns to be more informative
* Combines height columns into a single height in inches
* Converts date_stop to date data type
* Converts time_stop to time data type
* Recodes the values in categorical columns to be more informative
* Selects column subset for further analysis
* Removes the row of NAs that exists in the data

```{r, message=FALSE, include=FALSE}
# Read in data
stop_frisk_df = 
  # Read in data from internet
  GET("https://www1.nyc.gov/assets/nypd/downloads/excel/analysis_and_planning/stop-question-frisk/sqf-2016.csv") %>% 
  content("parsed") %>% 
  
  # Clean and fix names of columns
  janitor::clean_names() %>% 
  rename(
    precinct = pct,
    date_stop = datestop,
    time_stop = timestop,
    stop_in_out = inout,
    obs_time_min = perobs,
    stop_time_min = perstop,
    arst_made = arstmade,
    off_in_unif = offunif,
    hair_col = haircolr,
    eye_col = eyecolor,
    other_feature = othfeatr,
    boro = city
  )  %>% 
  mutate(
    # Combine height columns
    height_inch = ht_feet * 12 + ht_inch,
    # Convert date to proper format
    date_stop = mdy(date_stop),
    # Convert time to proper format
    time_stop = hm(time_stop / 100),
    # Recode to be more informative
    stop_in_out = recode(stop_in_out, "I" = "inside", "O" = "outside"),
    race = recode(
      race, 
      "A" = "asian/pacific islander", 
      "B" = "black", 
      "I" = "american indian/alaska native",
      "P" = "black-hispanic",
      "Q" = "white-hispanic",
      "W" = "white",
      "U" = "unknown",
      "Z" = "other"
    ),
    hair_col = recode(
      hair_col,
      "BA" = "bald",
      "BK" = "black",
      "BL" = "blond",
      "BR" = "brown",
      "DY" = "dyed",
      "FR" = "frosted",
      "GY" = "gray",
      "RA" = "red",
      "SN" = "sandy",
      "SP" = "salt and pepper",
      "WH" = "white",
      "XX" = "unknown",
      "ZZ" = "other",
    ),
    eye_col = recode(
      eye_col,
      "BK" = "black",
      "BL" = "blue",
      "BR" = "brown",
      "DF" = "different",
      "GR" = "green",
      "GY" = "gray",
      "HA" = "hazel",
      "MA" = "maroon",
      "PK" = "pink",
      "VI" = "violet",
      "XX" = "unknown",
      "Z" = "other",      
    ),
    build = recode(
      build,
      "H" = "heavy",
      "M" = "medium",
      "T" = "thin",
      "U" = "muscular",
      "Z" = "unknown"
    ),
    # change boro columns to lowercase for consistency
    boro = tolower(boro),
    # change character datatypes to numeric
    age = as.numeric(age),
    obs_time_min = as.numeric(obs_time_min),
    stop_time_min = as.numeric(stop_time_min)
  )  %>% 
  # select columns for further analysis
  select(precinct, date_stop, time_stop, stop_in_out, obs_time_min, stop_time_min, arst_made, off_in_unif, frisked, 
         searched, rf_vcrim, rf_othsw, rf_attir:ac_evasv, cs_furtv:cs_other, rf_knowl, sb_hdobj:sb_admis, rf_furt, 
         rf_bulg, sex, race, age, height_inch, weight:build, boro, xcoord, ycoord) %>% 
  # change all columns that have Y/N to 1/0
  mutate_at(vars(arst_made:rf_bulg), funs(recode(., "Y" = "1", "N" = "0"))) %>% 
  # change binary columns to numeric instead of character
  mutate_at(vars(arst_made:rf_bulg), funs(as.numeric(.))) %>% 
  # converts all character variables to factors (this does the same as the for loop)
  mutate_if(is.character, as.factor) %>% 
  # remove the single row of NAs
  filter(!is.na(build))
```


Final plots to keep

```{r}
# Number of stops per day
stop_frisk_df %>% 
  group_by(date_stop) %>% 
  summarize(
    count = n()
  ) %>% 
  ggplot(aes(x = date_stop, y = count)) + 
  geom_point() +
  geom_smooth(se = FALSE) +
  labs(
    title = "Number of stops per day in 2016",
    x = "Date of stop",
    y = "Number of stops"
  )

# Number of stops per month (broken down by boro) 
stop_frisk_df %>% 
  mutate(
    month_stop = month(date_stop),
    boro = str_to_title(boro)
  ) %>%   
  group_by(month_stop, boro) %>% 
  summarize(
    count = n()
  ) %>% 
  ggplot(aes(x = month_stop, y = count, color = boro)) + 
  geom_point() +
  geom_smooth(se = FALSE) +
  labs(
    title = "Number of stops per month by borough",
    x = "Month of stop",
    y = "Number of stops",
    color = "Borough"
  ) +
  scale_x_continuous(
    breaks = 1:12,
    labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
  )

# Number of stops per hour over the day
stop_frisk_df %>% 
  mutate(
    hour_stop = hour(time_stop),
    part_of_day = cut(hour_stop, breaks = c(0, 5, 12, 17, 21, 24), labels = c("Night", "Morning", "Afternoon", "Evening", "Night"), right = FALSE),
    part_of_day = fct_relevel(part_of_day, "Morning", "Afternoon", "Evening")
  ) %>% 
  select(hour_stop, part_of_day) %>% 
  group_by(hour_stop, part_of_day) %>% 
  summarize(
    count = n()
  ) %>% 
  ggplot(aes(x = hour_stop, y = count)) + 
  geom_bar(stat = "Identity", aes(fill = part_of_day)) +
  labs(
    title = "Number of stops per hour over the day",
    x = "Hour of stop",
    y = "Number of stops",
    fill = "Part of day"
  )

# Number of stops per day of week 
stop_frisk_df %>% 
  mutate(
    dow_stop = wday(date_stop, label = TRUE)
  ) %>% 
  group_by(dow_stop) %>% 
  summarize(
    count = n()
  ) %>% 
  ggplot(aes(x = dow_stop, y = count, fill = dow_stop)) + 
  geom_bar(stat = "Identity") +
  labs(
    title = "Number of stops per day of week",
    x = "Day of week",
    y = "Number of stops"
  ) +
  theme(legend.position = "none")
```



```{r, message=FALSE, eval=FALSE}
# Number of stops per day (KEEP)
stop_frisk_df %>% 
  group_by(date_stop) %>% 
  summarize(
    count = n()
  ) %>% 
  ggplot(aes(x = date_stop, y = count)) + 
  geom_point() +
  geom_smooth(se = FALSE)

# Number of stops per month
stop_frisk_df %>% 
  mutate(
    month_stop = factor(month(date_stop))
  ) %>% 
  filter(
    month_stop != is_null(month_stop)
  ) %>% 
  group_by(month_stop) %>% 
  summarize(
    count = n()
  ) %>% 
  ggplot(aes(x = month_stop, y = count)) + 
  geom_bar(stat = "Identity") 

# Number of stops per month (broken down by boro) (KEEP)
stop_frisk_df %>% 
  mutate(
    month_stop = month(date_stop)
  ) %>% 
  filter(
    month_stop != is_null(month_stop)
  ) %>%   
  group_by(month_stop, boro) %>% 
  summarize(
    count = n()
  ) %>% 
  ggplot(aes(x = month_stop, y = count, color = boro)) + 
  geom_point() +
  geom_smooth(se = FALSE) 

# Number of stops per hour over the day (KEEP)
stop_frisk_df %>% 
  mutate(
    hour_stop = hour(time_stop),
    part_of_day = cut(hour_stop, breaks = c(0, 5, 12, 17, 21, 24), labels = c("Night", "Morning", "Afternoon", "Evening", "Night"), right = FALSE)
  ) %>% 
  select(hour_stop, part_of_day) %>% 
  group_by(hour_stop, part_of_day) %>% 
  summarize(
    count = n()
  ) %>% 
  ggplot(aes(x = hour_stop, y = count)) + 
  geom_bar(stat = "Identity", aes(fill = part_of_day)) +
  geom_smooth(se = FALSE)

# Number of stops per hour over the day (broken down by boro)
stop_frisk_df %>% 
  mutate(
    hour_stop = hour(time_stop)
  ) %>% 
  group_by(hour_stop, boro) %>% 
  summarize(
    count = n()
  ) %>% 
  ggplot(aes(x = hour_stop, y = count, color = boro)) + 
  geom_point() +
  geom_smooth(se = FALSE)

# Number of stops per day of week (KEEP)
stop_frisk_df %>% 
  mutate(
    dow_stop = wday(date_stop, label = TRUE)
  ) %>% 
  group_by(dow_stop) %>% 
  summarize(
    count = n()
  ) %>% 
  ggplot(aes(x = dow_stop, y = count, fill = dow_stop)) + 
  geom_bar(stat = "Identity") 

# Number of stops per hour by day of week
stop_frisk_df %>% 
  mutate(
    dow_stop = wday(date_stop, label = TRUE),
    hour_stop = hour(time_stop)
  ) %>% 
  group_by(dow_stop, hour_stop) %>% 
  summarize(
    count = n()
  ) %>% 
  ggplot(aes(x = hour_stop, y = count, color = dow_stop)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

This code chunk looks at the number of people stopped, frisked, frisked & searched, and searched over a single day

```{r, message=FALSE, eval=FALSE}
stop_frisk_df %>% 
  filter(date_stop != is.na(date_stop)) %>% 
  mutate(
    hour_stop = hour(time_stop)
  ) %>% 
  group_by(hour_stop, frisked, searched) %>% 
  summarize(
    count = n()
  ) %>% 
  pivot_wider(
    names_from = frisked:searched,
    values_from = count
  ) %>% 
  rename(
    "stopped" = "0_0",
    "frisked" = "1_0",
    "searched" = "0_1",
    "frisk_and_search" = "1_1"
  ) %>% 
  mutate(
    searched = replace_na(searched, 0),
    stopped = replace_na(stopped, 0),
    frisked = replace_na(frisked, 0),
    frisk_and_search = replace_na(frisk_and_search, 0),
    stopped = stopped + searched + frisked + frisk_and_search
  ) %>% 
  ggplot(aes(x = hour_stop, y = stopped)) +
  geom_smooth(se = FALSE, color = 'red') +
  geom_smooth(aes(y = frisked, color = 'blue'), se = FALSE) +
  geom_smooth(aes(y = searched, color = 'green'), se = FALSE) +
  geom_smooth(aes(y = frisk_and_search, color = 'yellow'), se = FALSE) 
```

Animated 
```{r, eval=FALSE}
# Number of stops per hour over the day (broken down by boro)
stop_frisk_df %>% 
  mutate(
    hour_stop = hour(time_stop)
  ) %>% 
  group_by(hour_stop, boro) %>% 
  summarize(
    count = n()
  ) %>% 
  ggplot(aes(x = hour_stop, y = count, color = boro)) + 
  geom_point(alpha = 0.7, size = 2) +
  geom_smooth(se = FALSE, alpha = 0.7) +
  transition_states(
    boro,
    transition_length = 1,
    state_length = 1
  ) +
  enter_fade() +
  exit_shrink() +
  ggtitle('{closest_state}')
```

Animated 2 (option #1)
```{r, eval=FALSE}
# Number of stops per day
stop_frisk_df %>% 
  mutate(
    month = month(date_stop),
    day = day(date_stop)
  ) %>% 
  group_by(month, day) %>% 
  summarize(
    count = n()
  ) %>% 
  ggplot(aes(x = day, y = count, color = month)) + 
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(se = FALSE, alpha = 0.7) + 
  transition_states(
    month,
    transition_length = 2,
    state_length = 1
  ) +
  enter_fade() +
  exit_shrink() +
  ggtitle('Month: {closest_state}')
```

Animated 2 (option #2)
```{r,eval=FALSE}
# Number of stops per day
stop_frisk_df %>% 
  mutate(
    month = month(date_stop)
  ) %>% 
  group_by(month, date_stop) %>% 
  summarize(
    count = n()
  ) %>% 
  ggplot(aes(x = date_stop, y = count, color = month)) + 
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(se = FALSE, alpha = 0.7) + 
  transition_states(
    month,
    transition_length = 1,
    state_length = 1
  ) +
  enter_fade() +
  exit_shrink() +
  ggtitle('Month: {closest_state}') 
```

Animated
```{r, eval=FALSE}
stop_frisk_df %>% 
  mutate(
    dow_stop = wday(date_stop, label = TRUE),
    hour_stop = hour(time_stop)
  ) %>% 
  group_by(dow_stop, hour_stop) %>% 
  summarize(
    count = n()
  ) %>% 
  ggplot(aes(x = hour_stop, y = count, color = dow_stop)) + 
  geom_point() + 
  geom_smooth(se = FALSE) +
  transition_states(
    dow_stop,
    transition_length = 1,
    state_length = 1
  ) +
  enter_fade() +
  exit_shrink() +
  ggtitle('Day: {closest_state}') 
```








