---
title: "Plots"
author: "Keyanna Davis"
date: "11/30/2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(rvest)
library(httr)
library(lubridate)
library(plotly)
library(patchwork)
library(transformr)
library(gganimate)
library(transformr)
library(gifski)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

```{r}
# Read in data
stop_frisk_df = 
  # Read in data from internet
  GET("https://www1.nyc.gov/assets/nypd/downloads/excel/analysis_and_planning/stop-question-frisk/sqf-2016.csv") %>% 
  content("parsed") %>% 
  
  # Clean and fix names of columns
  janitor::clean_names() %>% 
  rename(
    precinct = pct,
    date_stop = datestop,
    time_stop = timestop,
    stop_in_out = inout,
    obs_time_min = perobs,
    stop_time_min = perstop,
    arst_made = arstmade,
    off_in_unif = offunif,
    hair_col = haircolr,
    eye_col = eyecolor,
    other_feature = othfeatr,
    boro = city
  )  %>% 
  mutate(
    # Combine height columns
    height_inch = ht_feet * 12 + ht_inch,
    # Convert date to proper format
    date_stop = mdy(date_stop),
    # Convert time to proper format
    time_stop = hm(time_stop / 100),
    # Recode to be more informative
    stop_in_out = recode(stop_in_out, "I" = "inside", "O" = "outside"),
    race = recode(
      race, 
      "A" = "other", 
      "B" = "black", 
      "I" = "other",
      "P" = "black-hispanic",
      "Q" = "white-hispanic",
      "W" = "white",
      "U" = "other",
      "Z" = "other"
    ),
    hair_col = recode(
      hair_col,
      "BA" = "bald",
      "BK" = "black",
      "BL" = "blond",
      "BR" = "brown",
      "DY" = "other",
      "FR" = "other",
      "GY" = "other",
      "RA" = "other",
      "SN" = "other",
      "SP" = "other",
      "WH" = "other",
      "XX" = "other",
      "ZZ" = "other",
    ),
    eye_col = recode(
      eye_col,
      "BK" = "black",
      "BL" = "blue",
      "BR" = "brown",
      "DF" = "other",
      "GR" = "other",
      "GY" = "other",
      "HA" = "other",
      "MA" = "other",
      "PK" = "other",
      "VI" = "other",
      "XX" = "other",
      "Z" = "other",      
    ),
    build = recode(
      build,
      "H" = "heavy",
      "M" = "medium",
      "T" = "thin",
      "U" = "other",
      "Z" = "unknown"
    ),
    # change boro columns to lowercase for consistency
    boro = tolower(boro),
    # change character datatypes to numeric
    age = as.numeric(age),
    obs_time_min = as.numeric(obs_time_min),
    stop_time_min = as.numeric(stop_time_min)
  )  %>% 
  # select columns for further analysis
  select(precinct, date_stop, time_stop, stop_in_out, obs_time_min, stop_time_min, arst_made, off_in_unif, frisked, 
         searched, rf_vcrim, rf_othsw, rf_attir:ac_evasv, cs_furtv:cs_other, rf_knowl, sb_hdobj:sb_admis, rf_furt, 
         rf_bulg, sex, race, age, height_inch, weight:build, boro, xcoord, ycoord) %>% 
  # change all columns that have Y/N to 1/0
  mutate_at(vars(arst_made:rf_bulg), funs(recode(., "Y" = "1", "N" = "0"))) %>% 
  # change binary columns to numeric instead of character
  mutate_at(vars(arst_made:rf_bulg), funs(as.numeric(.))) %>% 
  # converts all character variables to factors (this does the same as the for loop)
  mutate_if(is.character, as.factor) %>% 
  # remove the single row of NAs
  filter(!is.na(build))
```

PLOTS

PLOT 1: Percentage of reported stops/frisks by race 

```{r, plot1}
stop =
  stop_frisk_df %>% 
  drop_na() %>% 
  mutate(
    race = fct_infreq(race)
    ) %>% 
  ggplot(aes(race, fill = race))+
  geom_bar(aes(y=(..count..)/sum(..count..)),alpha = .5, position = "dodge")+
  labs(
    title = "Percentages of Reported Stops by Race",
    y= "percentage") +
    theme(legend.position = "none")

frisk =
  stop_frisk_df %>% 
  drop_na() %>% 
  mutate(
    race = fct_infreq(race)
    ) %>% 
  filter(frisked == '1') %>% 
  ggplot(aes(race, fill = race))+
  geom_bar(aes(y=(..count..)/sum(..count..)),alpha = .5, position = "dodge")+
  labs(
    title = "Percentages of Reported Stops Resulting in a Frisk by Race",
    y= "percentage") +
    theme(legend.position = "none")


(stop/frisk)
```

Plot 2: THe percentages of frisks by race

```{r, plot2}
  male=stop_frisk_df %>% 
  drop_na() %>% 
  mutate(
    race = fct_infreq(race)
    ) %>% 
  filter( 
    frisked == '1') %>% 
  filter(sex == "M") %>% 
  ggplot(aes(race, fill = sex))+
  geom_bar(aes(y=(..count..)/sum(..count..)),alpha = .5, position = "dodge")+
  labs(
    title = "Percentage of Reported Stops Resulting in a Frisk",
    y= "percentage")
 
 female = stop_frisk_df %>% 
  drop_na() %>% 
  mutate(
    race = fct_infreq(race)) %>% 
  filter( 
    frisked == '1') %>% 
  filter(sex == "F") %>% 
  ggplot(aes(race, fill = sex))+
  geom_bar(aes(y=(..count..)/sum(..count..)),alpha = .5, position = "dodge")+
  labs(
    title = "Percentage of Reported Stops Resulting in a Frisk",
    y= "percentage")
 
 (male/female)
```

PLOT 3: Percentage of Reported Stops Resulting in a Frisk by Age

```{r, plot3}
stop_age_group  =
  stop_frisk_df %>% 
  select(frisked, age) %>% 
  mutate(
    age_group = case_when(
      age < '18' ~ "18-",
      age >= '18' & age < '30' ~ "18-30",
      age >= '30' & age < '40' ~ "30s",
      age >= '40' & age < '50' ~ "40s",
      age >= '50' & age < '60' ~ "50s",
      age >= '60' ~"60+"
    )
    )
stop_age_group %>% 
   filter( 
    frisked == '1') %>% 
  ggplot(aes(age_group, fill = age_group))+
  geom_bar(aes(y=(..count..)/sum(..count..)),alpha = .5, position = "dodge")+
  labs(
    title = "Percentage of Reported Stops Resulting in a Frisk by Age",
    y= "percentage",
    x = "age") +
  theme(legend.position = "none")
```

PLOT 4: Percentage of Reported Stops Resulting in a Frisk by Borough

```{r, plot4}
stop_frisk_df %>% 
  ggplot(aes(boro, fill = boro))+
  geom_bar(aes(y=(..count..)/sum(..count..)),alpha = .5, position = "dodge")+
  labs(
    title = "Percentage of Reported Stops Resulting in a Frisk by Borough",
    y= "percentage",
    x = "borough") +
  theme(legend.position = "none")
```

PLOT 5: Total number of reasons for being stopped

```{r, plot5, eval = FALSE}
stop_frisk_df %>% 
  select(cs_objcs:cs_lkout, cs_cloth, cs_drgtr, cs_furtv, cs_vcrim:cs_other, frisked) %>% 
  pivot_longer(
    cs_objcs:cs_other,
    names_to = "reason_stopped",
    values_to = "stops"
  ) %>% 
  mutate(
    reason_stopped = recode(
    reason_stopped,
    "cs_objcs" = "carrying suspicious object",
    "cs_descr" = "fits a relevant description",
    "cs_casng" = "casing a victim or location",
    "cs_lkout" = "suspect acting as a lookout",
    "cs_cloth" = "wearing clothes commonly used in crimes",
    "cs_drgtr" = "actions indicative of drug transaction",
    "cs_furtv" = "furtive movements",
    "cs_vcrim" = "actions engaging in violent crime",
    "cs_bulge" = "suspcious bulge",
    "cs_other" = "other"
  )) %>% 
  filter(stops == 1) %>% 
  group_by(reason_stopped) %>% 
  summarize(total =n()) %>% 
  mutate(prop = total/sum(total)) %>% 
 plot_ly(x = ~reason_stopped, y= ~total, type = "bar" , color = ~reason_stopped) 
```

PLOT 6: The proportion of reasons being stopped by age group

```{r, plot6}
stop_age_group  =
  stop_frisk_df %>% 
  select(cs_objcs:cs_lkout, cs_cloth, cs_drgtr, cs_furtv, cs_vcrim:cs_other,age) %>% 
  mutate(
    age_group = case_when(
      age < '18' ~ "18-",
      age >= '18' & age < '30' ~ "18-30",
      age >= '30' & age < '40' ~ "30s",
      age >= '40' & age < '50' ~ "40s",
      age >= '50' & age < '60' ~ "50s",
      age >= '60' ~"60+"
    )
    ) %>% 
   pivot_longer(
    cs_objcs:cs_other,
    names_to = "reason_stopped",
    values_to = "stops"
  ) %>% 
  mutate(
    reason_stopped = recode(
    reason_stopped,
    "cs_objcs" = "carrying suspicious object",
    "cs_descr" = "fits a relevant description",
    "cs_casng" = "casing a victim or location",
    "cs_lkout" = "suspect acting as a lookout",
    "cs_cloth" = "wearing clothes commonly used in crimes",
    "cs_drgtr" = "actions indicative of drug transaction",
    "cs_furtv" = "furtive movements",
    "cs_vcrim" = "actions engaging in violent crime",
    "cs_bulge" = "suspcious bulge",
    "cs_other" = "other"
  )) %>% 
  filter(stops == 1) %>% 
  group_by(reason_stopped, age_group) %>% 
  summarize(total =n()) %>% 
  mutate(prob = total/sum(total)) %>% 
  drop_na

stop_age_group %>% 
  plot_ly(x = ~age_group, y= ~prob, type = "bar" , color = ~reason_stopped) 

age_dist = stop_frisk_df %>% 
  select(cs_objcs:cs_lkout, cs_cloth, cs_drgtr, cs_furtv, cs_vcrim:cs_other,age, race) %>% 
    ggplot(aes(x= age))+
   geom_density(aes(fill=race), alpha = .4)
ggplotly(age_dist)
```

PLOT 6.1: The total count of reasons being stopped by age group

```{r, plot6.1}
stop_age_group %>% 
  plot_ly(x = ~age_group, y= ~total, type = "bar" , color = ~reason_stopped)
```


PLOT 7: The proportion of reasons being stopped by sex

```{r, plot7}
stops_by_sex = 
  stop_frisk_df %>% 
  select(cs_objcs:cs_lkout, cs_cloth, cs_drgtr, cs_furtv, cs_vcrim:cs_other, sex) %>% 
  pivot_longer(
    cs_objcs:cs_other,
    names_to = "reason_stopped",
    values_to = "stops"
  ) %>% 
   mutate(
     sex = recode(
    sex,
    "F" = "female",
    "M" = "male",
    "Z" = "other"
  ), 
  reason_stopped = recode(
    reason_stopped,
    "cs_objcs" = "carrying suspicious object",
    "cs_descr" = "fits a relevant description",
    "cs_casng" = "casing a victim or location",
    "cs_lkout" = "suspect acting as a lookout",
    "cs_cloth" = "wearing clothes commonly used in crimes",
    "cs_drgtr" = "actions indicative of drug transaction",
    "cs_furtv" = "furtive movements",
    "cs_vcrim" = "actions engaging in violent crime",
    "cs_bulge" = "suspcious bulge",
    "cs_other" = "other"
  )) %>% 
   filter(stops == 1) %>% 
  group_by(reason_stopped, sex) %>% 
  summarize(total =n()) %>%
  mutate(prop = total/sum(total)) 

stops_by_sex %>% 
  plot_ly(x = ~sex, y= ~prop, type = "bar" , color = ~reason_stopped)

```

PLOT 7.1: The total count of reasons being stopped by sex

```{r, plot7.1}
stops_by_sex %>% 
  plot_ly(x = ~sex, y = ~total, type = "bar", color = ~reason_stopped)
```

PLOT 8: The porportion of reasons being stopped by race

```{r, plot8}
stop_by_race = 
  stop_frisk_df %>% 
  select(cs_objcs:cs_lkout, cs_cloth, cs_drgtr, cs_furtv, cs_vcrim:cs_other, race) %>% 
  pivot_longer(
    cs_objcs:cs_other,
    names_to = "reason_stopped",
    values_to = "stops"
  ) %>% mutate(
    reason_stopped = recode(
    reason_stopped,
    "cs_objcs" = "carrying suspicious object",
    "cs_descr" = "fits a relevant description",
    "cs_casng" = "casing a victim or location",
    "cs_lkout" = "suspect acting as a lookout",
    "cs_cloth" = "wearing clothes commonly used in crimes",
    "cs_drgtr" = "actions indicative of drug transaction",
    "cs_furtv" = "furtive movements",
    "cs_vcrim" = "actions engaging in violent crime",
    "cs_bulge" = "suspcious bulge",
    "cs_other" = "other"
  )) %>% 
   filter(stops == 1) %>% 
  group_by(reason_stopped, race) %>% 
  summarize(total =n()) %>% 
  mutate(prop = total/sum(total)) 

stop_by_race %>% 
  plot_ly(x = ~race, y= ~prop, type = "bar" , color = ~reason_stopped)
```

PLOT 8.1: The total count of reasons being stopped by race

```{r, plot8.1}
stop_by_race %>% 
  plot_ly(x = ~race, y= ~total, type = "bar" , color = ~reason_stopped)
```


PLOT 9: Arrest porportions of reasons being stopped 

```{r, plot9}
stops_arrest =
stop_frisk_df %>% 
  select(cs_objcs:cs_lkout, cs_cloth, cs_drgtr, cs_furtv, cs_vcrim:cs_other, arst_made) %>% 
  pivot_longer(
    cs_objcs:cs_other,
    names_to = "reason_stopped",
    values_to = "stops"
  ) %>% mutate(
    reason_stopped = recode(
    reason_stopped,
    "cs_objcs" = "carrying suspicious object",
    "cs_descr" = "fits a relevant description",
    "cs_casng" = "casing a victim or location",
    "cs_lkout" = "suspect acting as a lookout",
    "cs_cloth" = "wearing clothes commonly used in crimes",
    "cs_drgtr" = "actions indicative of drug transaction",
    "cs_furtv" = "furtive movements",
    "cs_vcrim" = "actions engaging in violent crime",
    "cs_bulge" = "suspcious bulge",
    "cs_other" = "other"
  )) %>% 
  mutate(
    arst_made = recode(
      arst_made,
      '0' = "not arrested",
      '1' = "arrested"
    )
  ) %>% 
   filter(stops == 1) %>% 
  group_by(reason_stopped, arst_made) %>% 
  summarize(total =n()) %>% 
  mutate( prob = total/sum(total))

stops_arrest %>% 
  plot_ly(x = ~arst_made, y= ~prob, type = "bar" , color = ~reason_stopped)
```

PLOT 9.1: The total count of arrest by reasons being stopped

```{r, plot9.1}
stops_arrest %>% 
  plot_ly(x = ~arst_made, y= ~total, type = "bar" , color = ~reason_stopped)
```

PLOT 10: The porportion of reasons being stopped by borough

```{r, plot10}
stop_by_boro = 
  stop_frisk_df %>% 
  select(cs_objcs:cs_lkout, cs_cloth, cs_drgtr, cs_furtv, cs_vcrim:cs_other, boro) %>% 
  pivot_longer(
    cs_objcs:cs_other,
    names_to = "reason_stopped",
    values_to = "stops"
  ) %>% mutate(
    reason_stopped = recode(
    reason_stopped,
    "cs_objcs" = "carrying suspicious object",
    "cs_descr" = "fits a relevant description",
    "cs_casng" = "casing a victim or location",
    "cs_lkout" = "suspect acting as a lookout",
    "cs_cloth" = "wearing clothes commonly used in crimes",
    "cs_drgtr" = "actions indicative of drug transaction",
    "cs_furtv" = "furtive movements",
    "cs_vcrim" = "actions engaging in violent crime",
    "cs_bulge" = "suspcious bulge",
    "cs_other" = "other"
  )) %>% 
   filter(stops == 1) %>% 
  group_by(reason_stopped, boro) %>% 
  summarize(total =n()) %>% 
  mutate( prob = total/sum(total))

stop_by_boro %>% 
  plot_ly(x = ~boro, y= ~prob, type = "bar" , color = ~reason_stopped)

stop_by_boro %>% 
  plot_ly(x = ~boro, y= ~total, type = "bar" , color = ~reason_stopped)
```

PLOT 10.1: The total count of reasons being stopped by borough

```{r, plot10.1}
stop_by_boro %>% 
  plot_ly(x = ~boro, y= ~total, type = "bar" , color = ~reason_stopped)
```



PLOT 11: Number of stops per day

```{r, plot11}
stop_frisk_df %>% 
  group_by(date_stop) %>% 
  summarize(
    count = n()
  ) %>% 
  ggplot(aes(x = date_stop, y = count)) + 
  geom_point() +
  geom_smooth(se = FALSE)
```

PLOT 12: Number of stops per month

```{r, plot12}
stop_frisk_df %>% 
  mutate(
    month_stop = factor(month(date_stop))
  ) %>% 
  filter(
    month_stop != is_null(month_stop)
  ) %>% 
  group_by(month_stop) %>% 
  summarize(
    count = n()
  ) %>% 
  ggplot(aes(x = month_stop, y = count)) + 
  geom_bar(stat = "Identity")
```

PLOT 13: Number of stops per day (broken down by boro)

```{r, plot13}
stop_frisk_df %>% 
  mutate(
    month_stop = month(date_stop)
  ) %>% 
  filter(
    month_stop != is_null(month_stop)
  ) %>%   
  group_by(month_stop, boro) %>% 
  summarize(
    count = n()
  ) %>% 
  ggplot(aes(x = month_stop, y = count, color = boro)) + 
  geom_point() +
  geom_smooth(se = FALSE)
```

PLOT 14: Number of stops per hour over the day

```{r, plot14}
stop_frisk_df %>% 
  mutate(
    hour_stop = hour(time_stop),
    part_of_day = cut(hour_stop, breaks = c(0, 5, 12, 17, 21, 24), labels = c("Night", "Morning", "Afternoon", "Evening", "Night"), right = FALSE)
  ) %>% 
  select(hour_stop, part_of_day) %>% 
  group_by(hour_stop, part_of_day) %>% 
  summarize(
    count = n()
  ) %>% 
  ggplot(aes(x = hour_stop, y = count)) + 
  geom_bar(stat = "Identity", aes(fill = part_of_day)) +
  geom_smooth(se = FALSE)

```

PLOT 15: Number of stops per hour over the day (broken down by boro)

```{r, plot15}
stop_frisk_df %>% 
  mutate(
    hour_stop = hour(time_stop)
  ) %>% 
  group_by(hour_stop, boro) %>% 
  summarize(
    count = n()
  ) %>% 
  ggplot(aes(x = hour_stop, y = count, color = boro)) + 
  geom_point() +
  geom_smooth(se = FALSE)
```

PLOT 16: Number of stops per day of week

```{r, plot16}
stop_frisk_df %>% 
  mutate(
    dow_stop = wday(date_stop, label = TRUE)
  ) %>% 
  group_by(dow_stop) %>% 
  summarize(
    count = n()
  ) %>% 
  ggplot(aes(x = dow_stop, y = count, fill = dow_stop)) + 
  geom_bar(stat = "Identity") 
```

PLOT 17: Number of stops per hour by day of week

```{r, plot17}
stop_frisk_df %>% 
  mutate(
    dow_stop = wday(date_stop, label = TRUE),
    hour_stop = hour(time_stop)
  ) %>% 
  group_by(dow_stop, hour_stop) %>% 
  summarize(
    count = n()
  ) %>% 
  ggplot(aes(x = hour_stop, y = count, color = dow_stop)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

PLOT 18: This code chunk looks at the number of people stopped, frisked, frisked & searched, and searched over a single day

```{r}
stop_frisk_df %>% 
  filter(date_stop != is.na(date_stop)) %>% 
  mutate(
    hour_stop = hour(time_stop)
  ) %>% 
  group_by(hour_stop, frisked, searched) %>% 
  summarize(
    count = n()
  ) %>% 
  pivot_wider(
    names_from = frisked:searched,
    values_from = count
  ) %>% 
  rename(
    "stopped" = "0_0",
    "frisked" = "1_0",
    "searched" = "0_1",
    "frisk_and_search" = "1_1"
  ) %>% 
  mutate(
    searched = replace_na(searched, 0),
    stopped = replace_na(stopped, 0),
    frisked = replace_na(frisked, 0),
    frisk_and_search = replace_na(frisk_and_search, 0),
    stopped = stopped + searched + frisked + frisk_and_search
  ) %>% 
  ggplot(aes(x = hour_stop, y = stopped)) +
  geom_smooth(se = FALSE, color = 'red') +
  geom_smooth(aes(y = frisked, color = 'blue'), se = FALSE) +
  geom_smooth(aes(y = searched, color = 'green'), se = FALSE) +
  geom_smooth(aes(y = frisk_and_search, color = 'yellow'), se = FALSE)
```
